---
title: "203 a4"
author: "Ruoyu Wang"
date: "2019/11/6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, message = FALSE, warnings = FALSE)
```

```{r}
library(tidyverse)
library(janitor)
library(wesanderson) ### color palette package

# Change all the -99999 values into NA and clean all the column names in to snake_case.
lob_rdata <- readr::read_csv("lobster_abundance_sbc_lter.csv",
                             na = "-99999") %>%
  clean_names()

# Make the raw data into tidy format by uncounting the observations to separeted rows.
lob_tidy <- lob_rdata %>% 
  tidyr::uncount(count)
```

# Introduction
Arroyo Quemado  Carpinteria  Isla Vista  Mohawk  Naples
	 
# Data and Methods 

# Results

Results A: Visually explore changes in annual lobster abundance (counts) by site. After grouping observations to find lobster counts for each year, create a finalized data visualization (no modeling/stats needed) showing changes in annual lobster abundance at the five sites over time.
```{r}
# Count the lobster abundance by site and year.
lob_year <- lob_tidy %>% 
  count(year, site)

# Show A
ggplot(data = lob_year, aes(x = year, y = n)) +
#  geom_col(aes(fill = site)) +
  geom_line(aes(color = site), 
            size = 1, 
            show.legend = FALSE) +
  geom_point(aes(color = site, shape = site), 
             size = 2.5, 
             show.legend = FALSE) +
  scale_shape_manual(values = c(15, 16,17, 18, 25)) +
  annotate("text", 
           x = 2018.5:2018.5:2018.5:2018.5:2018.5, 
           y = c(54, 343, 946, 164, 298), 
           label = c("AQUE", "CARP", "IVEE", "MOHK", "NAPL"),
           color = wes_palette("Darjeeling2", n = 5))+
  scale_color_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()
```

Results B. Visually explore lobster size distribution shifts by comparing lobster sizes in 2012 and 2018. Visually explore the size distributions for lobsters observed at the 5 different sites in 2012, and in 2018 (i.e. you will have a total of 10 size distributions shown: for each of the 5 sites, one for 2012 and one for 2018).

```{r}
lob_compare <- lob_tidy %>% 
  filter(year %in% c(2012, 2018)) %>% 
  select(year, site, size_mm)

# 10 histograms
ggplot(data = lob_compare, aes(x = size_mm)) +
  geom_histogram(aes(fill = site), show.legend = FALSE) +
  facet_grid(year~site, scales = "free")+
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()

# 10 qqplots
ggplot(data = lob_compare) +
  geom_qq(aes(sample = size_mm, color = site), show.legend = FALSE) +
  facet_grid(year~site, scales = "free")+
  scale_color_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()

# change facets' names
```

Results C. Compare mean lobster sizes at MPA vs. non-MPA sites in 2012 and 2018. Here, for each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters.

MPAs: NAPL, IVEE
Non-MPAs: AQUE, CARP, MOHK
Naples and Isla Vista, both established as MPAs on 2012-01-01.

```{r}
# Get the right category to different sites
lob_mpa <- lob_compare %>% 
    mutate(category = ifelse(site %in% c("NAPL", "IVEE"), "mpa", "non_mpa"))

# calculate the suammary table for the MPA group and non-MPA group
lob_mpa_summary <- lob_mpa %>% 
  group_by(category) %>% 
  summarize(
    site_mean = mean(size_mm, na.rm = TRUE),
    site_median = median(size_mm, na.rm = TRUE),
    site_sd = sd(size_mm, na.rm = TRUE),
    site_n = n()
  )
```

watch the data
```{r}
ggplot(data = lob_mpa, aes(x = size_mm, color = category, fill = category)) +
  geom_density(alpha = .3,show.legend = FALSE)+
  geom_vline(data = lob_mpa_summary, aes(xintercept = site_mean,
                                 colour = category),
             linetype = "dashed",
             size = 1
             ,show.legend = FALSE) +
  scale_y_discrete(expand = c(0,0)) +
  annotate("text", x = c(85, 62), y = .034:.034, label = c("MPA", "Non-MPA")) +
  theme_minimal()
```

two sample, two sided unpaired indenpendent t-test
null hypothesis: non-MPA and MPA have the same size mean
alternative hypothesis: the size mean in two groups are not the same

```{r}

```


# Summary

# References
