---
title: "203 a4"
author: "Ruoyu Wang"
date: "2019/11/6"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warnings = FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, message = FALSE, warnings = FALSE)
```

```{r}
library(tidyverse)
library(janitor)
library(wesanderson) ### color palette package
library(effsize)

# Change all the -99999 values into NA and clean all the column names in to snake_case.
lob_rdata <- readr::read_csv("lobster_abundance_sbc_lter.csv",
                             na = "-99999") %>%
  clean_names()

# Make the raw data into tidy format by uncounting the observations to separeted rows.
lob_tidy <- lob_rdata %>% 
  tidyr::uncount(count)
```

# Introduction
Arroyo Quemado  Carpinteria  Isla Vista  Mohawk  Naples
	 
# Data and Methods 

# Results

#### Results A: Visually explore changes in annual lobster abundance (counts) by site. After grouping observations to find lobster counts for each year, create a finalized data visualization (no modeling/stats needed) showing changes in annual lobster abundance at the five sites over time.
```{r}
# Count the lobster abundance by site and year.
lob_year <- lob_tidy %>% 
  count(year, site)

# Show A
ggplot(data = lob_year, aes(x = year, y = n)) +
#  geom_col(aes(fill = site)) +
  geom_line(aes(color = site), 
            size = 1, 
            show.legend = FALSE) +
  geom_point(aes(color = site, shape = site), 
             size = 2.5, 
             show.legend = FALSE) +
  scale_shape_manual(values = c(15, 16,17, 18, 25)) +
  annotate("text", 
           x = 2018.5:2018.5:2018.5:2018.5:2018.5, 
           y = c(54, 343, 946, 164, 298), 
           label = c("AQUE", "CARP", "IVEE", "MOHK", "NAPL"),
           color = wes_palette("Darjeeling2", n = 5))+
  scale_color_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()
```

##### Results B. Visually explore lobster size distribution shifts by comparing lobster sizes in 2012 and 2018. Visually explore the size distributions for lobsters observed at the 5 different sites in 2012, and in 2018 (i.e. you will have a total of 10 size distributions shown: for each of the 5 sites, one for 2012 and one for 2018).

```{r}
lob_compare_site <- lob_tidy %>% 
  filter(year %in% c(2012, 2018)) %>% 
  select(year, site, size_mm)

# 10 histograms
ggplot(data = lob_compare_site, aes(x = size_mm)) +
  geom_histogram(aes(fill = site), show.legend = FALSE) +
  facet_grid(year~site, scales = "free")+
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()

# 10 qqplots
ggplot(data = lob_compare_site) +
  geom_qq(aes(sample = size_mm, color = site), show.legend = FALSE) +
  facet_grid(year~site, scales = "free")+
  scale_color_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()

# change facets' names
```

#### Results C. Compare mean lobster sizes at MPA vs. non-MPA sites in 2012 and 2018. Here, for each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters.

MPAs: NAPL, IVEE
Non-MPAs: AQUE, CARP, MOHK
Naples and Isla Vista, both established as MPAs on 2012-01-01.

```{r}
# Get the right category to different sites
lob_tdata <- lob_compare_site %>% 
  mutate(category = ifelse(site %in% c("NAPL", "IVEE"), "mpa", "non_mpa"))

# calculate the suammary table for the MPA group and non-MPA group
lob_tsummary <- lob_tdata %>%
  group_by(category, year) %>%
  summarize(
    mean = mean(size_mm, na.rm = TRUE),
    median = median(size_mm, na.rm = TRUE),
    sd = sd(size_mm, na.rm = TRUE),
    n = n()
  )
```

watch the data
```{r}
ggplot(data = lob_tdata, aes(x = size_mm, color = category, fill = category)) +
  geom_density(alpha = .3,show.legend = FALSE)+
  facet_wrap(~year)+
  geom_vline(data = lob_tsummary, aes(xintercept = mean,
                                 colour = category),
             linetype = "dashed",
             size = 1,
             show.legend = FALSE) +
  scale_y_discrete(expand = c(0,0)) +
 # annotate("text", x = c(85, 62), y = .034:.034, label = c("MPA", "Non-MPA")) +
  theme_minimal()

ggplot(data = lob_tdata, aes(x = size_mm, fill = year, color = year)) +
  geom_density(alpha = .5,show.legend = FALSE)+
  facet_wrap(~category)+
  geom_vline(data = lob_tsummary, aes(xintercept = mean,
                                 colour = year),
             linetype = "dashed",
             size = 1,
             show.legend = FALSE) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 2)) +
 # annotate("text", x = c(85, 62), y = .034:.034, label = c("MPA", "Non-MPA")) +
  theme_minimal()
```

1. in 2012, size mean are the same in MPA and non-MPA sites
two sample, two sided unpaired indenpendent t-test
null hypothesis: non-MPA and MPA have the same size mean
alternative hypothesis: the size mean in two groups are not the same
```{r}
lob_2012 <- lob_tdata %>% 
  filter(year == 2012)

# ggplot(data = lob_mpa, aes(x = size_mm, fill = category, color = category)) +
#   geom_histogram(alpha = .5,show.legend = FALSE)+
#   geom_vline(data = lob_mpa_summary, aes(xintercept = site_mean,
#                                  colour = year),
#              linetype = "dashed",
#              size = 1,
#              show.legend = FALSE) +
#   scale_y_discrete(expand = c(0,0)) +
#   scale_fill_manual(values = wes_palette("Darjeeling2", n = 2)) +
#  # annotate("text", x = c(85, 62), y = .034:.034, label = c("MPA", "Non-MPA")) +
#   theme_minimal()

y2012_ttest <- t.test(data = lob_2012, size_mm~category)
y2012_d <- effsize::cohen.d(data = lob_2012, size_mm~category, na.rm = TRUE)
```

2. in 2018, size mean are the same in MPA and non-MPA sites
two sample, two sided unpaired indenpendent t-test
null hypothesis: non-MPA and MPA have the same size mean
alternative hypothesis: the size mean in two groups are not the same

```{r}
lob_2018 <- lob_tdata %>% 
  filter(year == 2018)

y2018_ttest <- t.test(data = lob_2018, size_mm~category)
y2018_d <- effsize::cohen.d(size_mm~category,data = lob_2018,na.rm = TRUE)
```

3. For MPA, mean size are the same in year 2012 and 2018
two sample, two sided unpaired indenpendent t-test
null hypothesis: 2012 and 2018 have the same size mean
alternative hypothesis: the size mean in two groups are not the same

```{r, message = FALSE, warnings = FALSE}
lob_mpa <- lob_tdata %>% 
  filter(category == "mpa")

mpa_ttest <- t.test(data = lob_mpa, size_mm~year)
mpa_d <- effsize::cohen.d(size_mm~year, data = lob_mpa, na.rm = TRUE)

```

4. For non-MPA, mean size are the same in year 2012 and 2018
two sample, two sided unpaired indenpendent t-test
null hypothesis: 2012 and 2018 have the same size mean
alternative hypothesis: the size mean in two groups are not the same

```{r, message = FALSE, warnings = FALSE}
lob_nonmpa <- lob_tdata %>% 
  filter(category == "non_mpa")

nonmpa_ttest <- t.test(data = lob_nonmpa, size_mm~year)
nonmpa_d <- effsize::cohen.d(data = lob_nonmpa, size_mm~year, na.rm = TRUE)
```
Cohen's *d* effecet size is calculated by: 

$$d = \frac{Mean_{MPA} - Mean_{non-MPA}}{SD_{pooled}}$$

where the pooled standard deviation for both groups, SD~pooled~ is calculated by: 

$$SD_{pooled}=\sqrt{\frac{SD_{MPA}^2 + SD_{non-MPA}^2}{2}}$$



**Example statement of test outcome:** 

"Mean size of lobster (mm) measured in MPA sites' samples (`r round(lob_tsummary$mean[1], 2)` $\pm$ `r round(lob_tsummary$sd[1], 2)`, n = `r lob_tsummary$n[1]`) differed significantly from lobster size in non MPA sites (`r round(lob_tsummary$mean[2], 2)` $\pm$ `r round(lob_tsummary$sd[2], 2)`, n = `r lob_tsummary$n[2]`) by a unpaired two-sample t-test (t(`r round(mpa_ttest$parameter, 2)`) = `r round(mpa_ttest$statistic, 2)`, *p* < 0.001)."

# Summary

# References
