---
output: html_document
---

# California Spiny Lobster Abindance Differences between Marine Protect Area and non protected area and time

#### Ruoyu Wang

#### 11/6/2019

```{r setup, include=FALSE, message = FALSE, warnings = FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, message = FALSE, warnings = FALSE, fig.align = "center", comment = FALSE)
```

```{r}
# Attach packages
library(tidyverse)
library(janitor)
library(wesanderson) ### color palette package
library(effsize)

# Tidy up the data:
# 1. Tell R all the -99999 values are actually NAs. Then clean all the column names into snake_case.
lob_rdata <- readr::read_csv("lobster_abundance_sbc_lter.csv",
                             na = "-99999") %>%
  clean_names()

# 2. Make the raw data into tidy format by uncounting the observations to separeted rows.
lob_tidy <- lob_rdata %>% 
  tidyr::uncount(count)
```


### Introduction
Arroyo Quemado  Carpinteria  Isla Vista  Mohawk  Naples
	 
	 
### Data and Methods 



### Results

#### 1. Changes in annual lobster abundance by site.
```{r annual change}
# Count the lobster amount by site and year.
lob_year <- lob_tidy %>% 
  count(year, site)

# Show A
ggplot(data = lob_year, aes(x = year, y = n)) +
#  geom_col(aes(fill = site)) +
  geom_line(aes(color = site), 
            size = 1, 
            show.legend = FALSE) +
  geom_point(aes(color = site, shape = site), 
             size = 2.5, 
             show.legend = FALSE) +
  scale_shape_manual(values = c(15, 16,17, 18, 25)) +
  annotate("text", 
           x = 2018.5:2018.5:2018.5:2018.5:2018.5, 
           y = c(54, 343, 946, 164, 298), 
           label = c("AQUE", "CARP", "IVEE", "MOHK", "NAPL"),
           color = wes_palette("Darjeeling2", n = 5))+
  scale_color_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()
```
***Figure. 1*** *lobster abundance changes from 2012 to 2018 in four different sites.*


#### 2. Lobster size distribution observed at the 5 different sites in 2012 and 2018.

```{r}
lob_compare_site <- lob_tidy %>% 
  filter(year %in% c(2012, 2018)) %>% 
  select(year, site, size_mm)

site_labs <- c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")
names(site_labs) <- c("AQUE","CARP","IVEE","MOHK", "NAPL")
# 10 histograms
ggplot(data = lob_compare_site, aes(x = size_mm)) +
  geom_histogram(aes(fill = site), show.legend = FALSE) +
  labs(x = "Spiny lobster size (mm)", y = "Abundance in count")+
  facet_grid(year~site, scales = "free",
             labeller = labeller(site = site_labs))+
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal() 
#, facet = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")
```

***Figure. 1*** *lobster abundance changes from 2012 to 2018 in four different sites.*

```{r}
# 10 qqplots
ggplot(data = lob_compare_site) +
  geom_qq(aes(sample = size_mm, color = site), show.legend = FALSE) +
  facet_grid(year~site, scales = "free",
             labeller = labeller(site = site_labs))+
  scale_color_manual(values = wes_palette("Darjeeling2", n = 5)) +
  theme_minimal()
 
# change facets' names
```

***Figure. 1*** *lobster abundance changes from 2012 to 2018 in four different sites.*


#### 3. Compare mean lobster sizes at MPA vs. non-MPA sites in 2012 and 2018. Here, for each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters.

MPAs: NAPL, IVEE
Non-MPAs: AQUE, CARP, MOHK
Naples and Isla Vista, both established as MPAs on 2012-01-01.

```{r}
# Get the right category to different sites
lob_tdata <- lob_compare_site %>% 
  mutate(category = ifelse(site %in% c("NAPL", "IVEE"), "mpa", "non_mpa")) %>% 
  mutate(year = as.character(year))

# calculate the suammary table for the MPA group and non-MPA group
lob_tsummary <- lob_tdata %>%
  group_by(category, year) %>%
  summarize(
    mean = mean(size_mm, na.rm = TRUE),
    median = median(size_mm, na.rm = TRUE),
    sd = sd(size_mm, na.rm = TRUE),
    n = n()
  )
```

watch the data

```{r}
ggplot(data = lob_tdata, aes(x = size_mm, color = category, fill = category)) +
  geom_density(alpha = .3,show.legend = FALSE)+
  facet_wrap(~year)+
  geom_vline(data = lob_tsummary, aes(xintercept = mean,
                                 colour = category),
             linetype = "dashed",
             size = 1,
             show.legend = FALSE) +
  scale_fill_manual(values = c("#56B4E9", "#CC79A7"))+
  scale_color_manual(values = c("#56B4E9", "#CC79A7"))+
  scale_y_discrete(expand = c(0,0)) +
  theme_minimal()
```

***Figure. 1(a)*** *lobster abundance changes from 2012 to 2018 in four different sites.*

```{r}
ggplot(data = lob_tdata, aes(x = size_mm, fill = year, color = year)) +
  geom_density(alpha = .3,show.legend = FALSE)+
  facet_wrap(~category)+
  geom_vline(data = lob_tsummary, aes(xintercept = mean,
                                 colour = year),
             linetype = "dashed",
             size = 1,
             show.legend = FALSE) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_manual(values = c("#E69F00", "#009E73")) +
  scale_color_manual(values = c("#E69F00", "#009E73")) +
  #annotate("text", x = c(85, 62), y = .034:.034, label = c("MPA", "Non-MPA"), size = 0.5) +
  theme_minimal()
```

***Figure. 1(b)*** *lobster abundance changes from 2012 to 2018 in four different sites.*

1. in 2012, size mean are the same in MPA and non-MPA sites
two sample, two sided unpaired indenpendent t-test

```{r}
lob_2012 <- lob_tdata %>% 
  filter(year == 2012)

y2012_ttest <- t.test(data = lob_2012, size_mm~category)
y2012_d <- effsize::cohen.d(data = lob_2012, size_mm~category, na.rm = TRUE)
y2012_ttest
y2012_d 
```

In year 2012, mean size of lobster (mm) measured in MPA sites' samples (`r round(lob_tsummary$mean[1], 2)` $\pm$ `r round(lob_tsummary$sd[1], 2)`, n = `r lob_tsummary$n[1]`) differed significantly from lobster size in non MPA sites (`r round(lob_tsummary$mean[3], 2)` $\pm$ `r round(lob_tsummary$sd[3], 2)`, n = `r lob_tsummary$n[3]`) by a unpaired two-sample t-test (t(`r round(y2012_ttest$parameter, 2)`) = `r round(y2012_ttest$statistic, 2)`, *p*(`r round(y2012_ttest$p.value, 3)`) < 0.05). The differnece in mean size between the two groups is ..., with a ...effect size (Cohen's d = `r round(y2012_d$estimate, 2)`). Therefore, ...

2. in 2018, size mean are the same in MPA and non-MPA sites

two sample, two sided unpaired indenpendent t-test

```{r}
lob_2018 <- lob_tdata %>% 
  filter(year == 2018)

y2018_ttest <- t.test(data = lob_2018, size_mm~category)
y2018_d <- effsize::cohen.d(size_mm~category,data = lob_2018,na.rm = TRUE)
y2018_ttest
y2018_d
```

In year 2018, mean size of lobster (mm) measured in MPA sites' samples (`r round(lob_tsummary$mean[2], 2)` $\pm$ `r round(lob_tsummary$sd[2], 2)`, n = `r lob_tsummary$n[2]`) differed significantly from lobster size in non MPA sites (`r round(lob_tsummary$mean[4], 2)` $\pm$ `r round(lob_tsummary$sd[4], 2)`, n = `r lob_tsummary$n[4]`) by a unpaired two-sample t-test (t(`r round(y2018_ttest$parameter, 2)`) = `r round(y2018_ttest$statistic, 2)`, *p*< 0.001). The differnece in mean size between the two groups is ..., with a smalleffect size (Cohen's d = `r round(y2018_d$estimate, 2)`). Therefore, ...

3. For MPA, mean size are the same in year 2012 and 2018
two sample, two sided unpaired indenpendent t-test


```{r, message = FALSE, warnings = FALSE}
lob_mpa <- lob_tdata %>% 
  filter(category == "mpa") 
 
mpa_ttest <- t.test(data = lob_mpa, size_mm~year)
mpa_d <- effsize::cohen.d(size_mm~year, data = lob_mpa, na.rm = TRUE)
mpa_ttest
mpa_d
```

For MPA samples, mean size of lobster (mm) measured in year 2012 (`r round(lob_tsummary$mean[1], 2)` $\pm$ `r round(lob_tsummary$sd[1], 2)`, n = `r lob_tsummary$n[1]`) differed significantly from lobster size in year 2018 (`r round(lob_tsummary$mean[2], 2)` $\pm$ `r round(lob_tsummary$sd[2], 2)`, n = `r lob_tsummary$n[2]`) by a unpaired two-sample t-test (t(`r round(mpa_ttest$parameter, 2)`) = `r round(mpa_ttest$statistic, 2)`, *p*< 0.001). The differnece in mean size between the two groups is ..., with a large effect size (Cohen's d = `r round(mpa_d$estimate, 2)`). Therefore, ...

4. For non-MPA, mean size are the same in year 2012 and 2018
two sample, two sided unpaired indenpendent t-test


```{r, message = FALSE, warnings = FALSE}
lob_nonmpa <- lob_tdata %>% 
  filter(category == "non_mpa") 

nonmpa_ttest <- t.test(data = lob_nonmpa, size_mm~year)
nonmpa_d <- effsize::cohen.d(size_mm~year, data = lob_nonmpa)
nonmpa_ttest
nonmpa_d
```

For non-MPA samples, mean size of lobster (mm) measured in year 2012 (`r round(lob_tsummary$mean[3], 2)` $\pm$ `r round(lob_tsummary$sd[3], 2)`, n = `r lob_tsummary$n[3]`) does not different significantly from lobster size in year 2018 (`r round(lob_tsummary$mean[4], 2)` $\pm$ `r round(lob_tsummary$sd[4], 2)`, n = `r lob_tsummary$n[4]`) by a unpaired two-sample t-test (t(`r round(nonmpa_ttest$parameter, 2)`) = `r round(nonmpa_ttest$statistic, 2)`, *p* = `r round(nonmpa_ttest$p.value, 2)`). The differnece in mean size between the two groups is ..., with a negligible effect size (Cohen's d = `r round(nonmpa_d$estimate, 2)`). Therefore, ...




# Summary

# References
